Unsupervised Semantic Mapping for Healthcare Data Storage Schema

### 2021-06-05 ManuScript/IEEEAccess/ManuScript.tex
kappa upated
update

\begin{table}[]
	\centering
	\caption{Abbreviations}
	\label{tab:abbreviations}
	\begin{tabular}{ll}
		\hline
		Abbreviations              & Full Form                                               \\ \hline
		ICT                        & Information and Communication Technologies              \\
		HMIS                       & Healthcare Management Information Systems               \\
		CDSS                       & Clinical Decision Support Systems                       \\
		HL7                        & Health Level Seven International                        \\
		HIMSS                      & Healthcare Information and Management Systems Society   \\
		FHIR                       & Fast Healthcare Interoperability Resources              \\
		Snomed-CT                  & Systematized Nomenclature of Medicine—Clinical Terms    \\
		LOINC                      & Logical Observation Identifiers Names and Codes         \\
		MLM                        & Medical Logic Module                                    \\
		CIMI                       & Clinical Information Modeling Initiative                \\
		UHP                        & Ubiquitous Health Platform                              \\
		NLP                        & Natural Language Processing                             \\
		BERT                       & Bidirectional Encoder Representations from Transformers \\
		UMLS                       & Unified Medical Language System                         \\
		EMR 					   & Electronic Medical Records                              \\
		EHR						   & Electronic Health Records                               \\
		MCC                        & Matthews Correlation coefficient                        \\
		NLI                        & Natural Language Inference                              \\ \hline
	\end{tabular}
\end{table}

@@

Healthcare Standards such as HL7 - Fast Healthcare Interoperability Resources (FHIR), and openEHR provide the foundations for storing and communicating medical data, through the use of well defined protocols. While Systematized Nomenclature of Medicine—Clinical Terms (Snomed-CT) \cite{snomedct_url} and Logical Observation Identifiers Names and Codes (LOINC) \cite{loinc} provide a standard definition for clinical terminologies and laboratory tests, respectively. Similarly Medical Logic Module (MLM) provides a standardized way for expressing medical knowledge. However, the plethora of standards necessitates the creation of bridging standards that can resolve the heterogeniety between the medical standards. Substantial effort has gone into this endeavor with the Clinical Information Modeling Initiative (CIMI) \cite{CIMI2015} taking the lead in bridging the gap between HL7v3 and openEHR. Similarly, SNOMED CT and LOINC are working to resolve the redundancies between the two terminological standards since 2013. This healthcare interoperability solution follows a formal, albiet long process, which is greatly dependent on the human factor. However, the current healthcare scenario requires a quick solution to create a scaffolding of an interoperable bridge between various healthcare providers. It is also important to ensure that this scaffolding should be able to support the formal standardization processes of the future. In \cite{Satti2020}, we have presented the Ubiquitous Health Platform (UHP), which provides semantic reconciliation-on-read based data curation for resolving data interoperability between various schema. This methodology is based on the creation and management of schema maps, that can provide the framework for transforming a source schema into a target schema. 

@@

\begin{itemize}
	\item[$-$] Section \ref{relatedwork} presents the related work
	\item[$-$] Section \ref{methodology} contains the details of our methodology
	\item[$-$] Section \ref{experimentalSetup} provides the experimental setup
	\item[$-$] Section \ref{results} presents the results
	\item[$-$] Section 6 concludes the paper.
\end{itemize}

@@

\begin{figure*}[t!]
	\centering
	\includegraphics[scale=0.83]{evaluation}
	\caption{Performance evaluation of various models using MCC and $\kappa$ scores.}
	\label{fig:evaluation}
\end{figure*}

\section{Discussion}
In text classification, production and use of a well annotated corpus for supervised and semi-supervised learning is of utmost importance. The same is also useful for evaluating the performance of unsupervised learning techniques. However, in the real world the production and maintenance of these corpora is an expensive task, often requiring extensive human effort. While there are many factors, influencing this reality, one of the most critical is the perception and cognitive ability of an expert user to subjectively provide a label to an instance. This is also evident in the case of our annotated dataset, whereby the instances have been labeled by four human experts (two medical practioners and two computer science graduates), using their subjective knowledge. One example of this subjective classification, is evident in the raw form of the data instances labeled by ``Annotator3''. According to ``Annotator3'' the relationship between the term ``AdmissionId'' and ``ClinicalHistory'' is parent-child. While the selection of this label can be debated from an subjective view point, changing or removing it or any other label, from the (objective) view point of the computing methodology would be incorrect. As a result, the annotations were kept anonymous with random to as not to induce any biasedness. Hence, the annotated data in its original form becomes the basis of our threshold selection and model evaluation methodology. A side-effect of this choice is the lower scores of agreement between the (annotated data) true labels and predicted labels.

Throughout this research work, the choice of performance metrics used for threshold selection and model evaluations are also driven by the nature of the dataset. Even before annotation, the dataset is bias in favour of unrelated attributes. As established by the human experts and the machine learning model, for 254 attributes involved in 40698 possible pairs only a little over 300 similar instances are found. In these circumstances performance metrics, such as accuracy, precision, and recall are meaningless. These metrics are unable to account for the imbalanced datasets and provide an incorrect view of the classifier's accuracy. Instead metrics such as $F_1$, MCC, $\kappa$ can provide a true picture of the classifier's accuracy. These metrics are also well suited for evaluating multi-class classifiers, using one vs all or one vs one binarizations of the dataset, as well as consolidating the results into a single measure. In our experimentations, we also evaluated the area under the Receiver Operating Characteristic (ROC) curve and area under the Precision Recall curve, as a threshold selection metric. However these graphs are well suited for independent classes and in our case it is important to maximize the identification of similar attribute pairs, followed by related ones and finally the unrelated ones. This kind of dependent classification can then be used in practice to identify a small set of similar attribute pairs, which can be used to establish positive results.

\section{Conclusion}
Healthcare information systems are slowly moving towards a standardized approach, whereby the storage, exchange, and usage of health data, information, and knowledge is becoming interoperable. However, research and development of ubiquitous health management systems, still require manual interventions by human experts. These approaches require costly resource allocations, which is not possible for small and medium scaled healthcare service providers. In order to provide data interoperability services to these providers, and automating the process of generating maps between HMIS schema, which can match and transform patient data from a source provider to a target one, it is important to utilize state-of-the-art NLP techniques. Using transformer based pre-trained BERT models, we show the applicability of these techniques on creating the, as yet non-validated, schema maps. Resultantly, human effort can now focus on validating and verifying the schema maps, which can greatly accelerate the application of healthcare data interoperability, and provide a cohesive ecosystem for ubiquitous healthcare service delivery.
In future, we shall look towards developing a mechanism for schema map storage and evolution, along with expert validation to finally close this loop. We shall also work on re-training some of the existing pre-trained models, to improve the performance of schema map creation.

>>>>

\begin{table}[]
	\centering
	\caption{Abbreviations}
	\label{tab:abbreviations}
	\begin{tabular}{ll}
		\hline
		Abbreviations              & Full Form                                               \\ \hline
		ICT                        & Information and Communication Technologies              \\
		HMIS                       & Healthcare Management Information Systems               \\
		CDSS                       & Clinical Decision Support Systems                       \\
		HL7                        & Health Level Seven International                        \\
		HIMSS                      & Healthcare Information and Management Systems Society   \\
		FHIR                       & Fast Healthcare Interoperability Resources              \\
		Snomed-CT                  & Systematized Nomenclature of Medicine—Clinical Terms    \\
		LOINC                      & Logical Observation Identifiers Names and Codes         \\
		MLM                        & Medical Logic Module                                    \\
		openCIMI                       & Clinical Information Modeling Initiative                \\
		UHP                        & Ubiquitous Health Platform                              \\
		NLP                        & Natural Language Processing                             \\
		BERT                       & Bidirectional Encoder Representations from Transformers \\
		UMLS                       & Unified Medical Language System                         \\
		EMR 					   & Electronic Medical Records                              \\
		EHR						   & Electronic Health Records                               \\
		MCC                        & Matthews Correlation coefficient                        \\
		NLI                        & Natural Language Inference                              \\ \hline
	\end{tabular}
\end{table}

@@

Healthcare Standards such as HL7 - Fast Healthcare Interoperability Resources (FHIR), and openEHR provide the foundations for storing and communicating medical data, through the use of well defined protocols. While Systematized Nomenclature of Medicine—Clinical Terms (Snomed-CT) \cite{snomedct_url} and Logical Observation Identifiers Names and Codes (LOINC) \cite{loinc} provide a standard definition for clinical terminologies and laboratory tests, respectively. Similarly Medical Logic Module (MLM) provides a standardized way for expressing medical knowledge. However, the plethora of standards necessitates the creation of bridging standards that can resolve the heterogeniety between the medical standards. Substantial effort has gone into this endeavor with the Clinical Information Modeling Initiative (openCIMI) \cite{CIMI2015} taking the lead in bridging the gap between HL7v3 and openEHR. Similarly, SNOMED CT and LOINC are working to resolve the redundancies between the two terminological standards since 2013. This healthcare interoperability solution follows a formal, albiet long process, which is greatly dependent on the human factor. However, the current healthcare scenario requires a quick solution to create a scaffolding of an interoperable bridge between various healthcare providers. It is also important to ensure that this scaffolding should be able to support the formal standardization processes of the future. In \cite{Satti2020}, we have presented the Ubiquitous Health Platform (UHP), which provides semantic reconciliation-on-read based data curation for resolving data interoperability between various schema. This methodology is based on the creation and management of schema maps, that can provide the framework for transforming a source schema into a target schema. 

@@

\begin{itemize}
	\item[$-$] Section \ref{relatedwork} presents the related work
	\item[$-$] Section \ref{methodology} contains the details of our methodology
	\item[$-$] Section \ref{experimentalSetup} provides the experimental setup
	\item[$-$] Section \ref{results} presents the results
	\item[$-$] Section \ref{discussion} discusses the challenges and limitations of the presented methodology and its evaluations
	\item[$-$] Section \ref{conclusion} concludes the paper.
\end{itemize}

@@

\begin{figure*}[t!]
	\centering
	\includegraphics[scale=0.83]{evaluation}
	\caption{Performance evaluation of various models using MCC and kappa ($\kappa$) scores.}
	\label{fig:evaluation}
\end{figure*}

\section{Discussion}
\label{discussion}
In text classification, production and use of a well annotated corpus for supervised and semi-supervised learning is of utmost importance. The same is also useful for evaluating the performance of unsupervised learning techniques. However, in the real world the production and maintenance of these corpora is an expensive task, often requiring extensive human effort. While there are many factors, influencing this reality, one of the most critical is the perception and cognitive ability of an expert user to subjectively provide a label to an instance \cite{reidsma2008exploiting, velupillai2010towards}. This is also evident in the case of our annotated dataset, whereby the instances have been labeled by four human experts (two medical practioners and two computer science graduates), using their subjective knowledge. One example of this subjective classification, is evident in the raw form of the data instances labeled by ``Annotator3''. According to ``Annotator3'' the relationship between the term ``AdmissionId'' and ``ClinicalHistory'' is parent-child. While the selection of this label can be debated from an subjective view point, changing or removing it or any other label, from the (objective) view point of the computing methodology would be incorrect\cite{pavlick2019inherent}. As a result, the annotations were kept anonymous so as not to induce any bias. Thus, the complete annotated data in its original form became the basis of our threshold selection and model evaluation methodology. A mode based voting mechanism was then used to resolve the differences between the annotators. The consolidated true dataset was then formed based on the agreed upon label by atleast 3 annotators. As pointed out in \cite{reidsma2008exploiting}, the net effect of such a voting mechanism is an increase in the precision of the machine learning classifiers in lieu of, their accuracy. However, as shown in the results and discussed further on, due to the bias nature of our dataset, accuracy measure is replaced by MCC. The lower scores of agreement between the (annotated data) true labels and predicted labels, have thus been evaluated in a contextful and relative manner.

Throughout this research work, the choice of performance metrics used for threshold selection and model evaluations are also driven by the nature of the dataset. Even before annotation, the dataset is bias in favour of unrelated attributes. As established by the human experts and the machine learning model, for 254 attributes involved in 40698 possible pairs only a little over 300 similar instances are found. In these circumstances performance metrics, such as accuracy, precision, and recall are meaningless. These metrics are unable to account for the imbalanced datasets and provide an incorrect view of the classifier's accuracy. Instead metrics such as $F_1$, MCC, and $\kappa$ can provide a true picture of the classifier's accuracy. These metrics are also well suited for evaluating multi-class classifiers, using one vs all or one vs one binarizations of the dataset, as well as consolidating the results into a single measure. In our experimentation, we also evaluated the Area Under the Receiver Operating Characteristic (AUROC) curve and Area Under the Precision Recall (AUPR) curve, as a threshold selection metric. However these graphs are well suited for independent classes as shown in \cite{espindola2005extending}, and in our case it is important to maximize the identification of similar attribute pairs ($class_1$), followed by related ones ($class_{0.5}$) and finally the unrelated ones($class_0$). AUROC and AUPR were thus replaced with our current approach for threshold selection. The benefit of using this kind of dependent classification is its usefulness in practice to identify a small set of similar attribute pairs, which can be used to establish positive results.

\section{Conclusion}
\label{conclusion}
Healthcare information systems are slowly moving towards a standardized approach, whereby the storage, exchange, and usage of health data, information, and knowledge is becoming interoperable. However, research and development of ubiquitous health management systems, still require manual interventions by human experts. These approaches require costly resource allocations, which is not possible for small and medium scaled healthcare service providers. In order to provide data interoperability services to these providers, and automating the process of generating maps between HMIS schema, which can match and transform patient data from a source provider to a target one, it is important to utilize state-of-the-art NLP techniques. Using transformer based pre-trained BERT models, we show the applicability of these techniques on creating the, as yet non-validated, schema maps. Resultantly, human effort can now focus on validating and verifying the schema maps, which can greatly accelerate the application of healthcare data interoperability, and provide a cohesive ecosystem for ubiquitous healthcare service delivery.
In future, we shall look towards developing a mechanism for schema map storage and evolution, along with expert validation to finally close this loop. We shall also work on re-training some of the existing pre-trained models, to improve the performance of schema map creation.
